---
title: "Assignment 2"
author: "Molly F-F & Brittany O"
date: "2024-09-28"
output: html_document
---

```{r}
library(tidyverse)
library(gtrendsR)
library(censusapi)
library(dplyr)
```

#Pulling from APIs
Our first data source is the Google Trends API. Suppose we are interested in the search trends
for crime and loans in Illinois in the year 2020. We could find this using the following code:
```{r}
set.seed(1)
res <- gtrends(c("crime", "loans"),
               geo = "US-IL",
               time = "2020-01-01 2020-12-31",
               low_search_volume = TRUE)
plot(res)
```

Answer the following questions for the keywords “crime” and “loans”.

• Find the mean, median and variance of the search hits for the keywords.

CRIME:                     LOANS
Mean: 51.01887             Mean: 62.24528
Median: 51                 Median: 60
Var.: 67.36502             Var: 98.72714
```{r}
head(res$interest_over_time)
res_time <- as_tibble(res$interest_over_time)
glimpse(res_time)

res_time_w <- spread(res_time, key = keyword, value = hits)
head(res_time_w)

mean(res_time_w$crime) #51.01887
median(res_time_w$crime) #51
var(res_time_w$crime) #67.36502


mean(res_time_w$loans) #62.24528
median(res_time_w$loans) #60
var(res_time_w$loans) #98.72714
```

• Which cities (locations) have the highest search frequency for loans? Note that there
might be multiple rows for each city if there were hits for both “crime” and “loans” in
that city. It might be easier to answer this question if we had the search hits info for
both search terms in two separate variables. That is, each row would represent a unique
city.

- Long Lake, Peotone, Rosemont, East Saint Louis, and Ford Heights are the top 5 locations for 'loans' search hits. 
```{r}
head(res$interest_by_city)
res_city <- as_tibble(res$interest_by_city)
glimpse(res_city)

res_city_w <- spread(res_city, key = keyword, value = hits)
head(res_city_w)
loansbycity <-  res_city_w %>%
  group_by(location) %>%
  summarize(loans = sum(loans)) %>%
  arrange(desc(loans))

na.omit(loansbycity)
```

• Is there a relationship between the search intensities between the two keywords we used?
```{r}
res_both_w <- res_city_w %>% 
  drop_na() 
res_both_w
```

In the cities that have data for both search keywords, the search intensity for "loans" is greater than it is for "crime." We can also see looking in general in the state of Illinois that throughout 2020, with the exception of the spike in roughly June 2020, the search intensity for "loans" is greater than the search intensity for "crime."

###Repeat the above for keywords related to covid. Make sure you use multiple keywords like we did above. Try several different combinations and think carefully about words that might make sense within this context.
```{r}
covid.search1 <- gtrends(c("covid", "vaccine"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search1)
```
```{r}
covid.search2 <- gtrends(c("covid", "test"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search2)
```
```{r}
covid.search3 <- gtrends(c("covid", "lockdown"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search3)
```
```{r}
covid.search4 <- gtrends(c("covid", "mask"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search4)
```
```{r}
covid.search5 <- gtrends(c("covid", "distancing"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search5)
```
```{r}
covid.search6 <- gtrends(c("covid", "China"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search6)
```
```{r}
covid.search7 <- gtrends(c("covid", "trump"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search7)
```
```{r}
covid.search8 <- gtrends(c("covid", "biden"),
geo = "US-IL",
time = "2020-01-01 2020-12-31",
low_search_volume = TRUE)
plot(covid.search8)
```
```{r}
rm(covid.search1,covid.search2, covid.search3, covid.search5, covid.search6, covid.search7, covid.search8)

head(covid.search4$interest_over_time)
covid_time <- as_tibble(covid.search4$interest_over_time)
glimpse(covid_time)

covid_time_w <- spread(covid_time, key = keyword, value = hits)
head(covid_time_w)
glimpse(covid_time_w)

covid_time_w$covid <- as.numeric(covid_time_w$covid)
covid_time_w$mask <- as.numeric(covid_time_w$mask)
glimpse(covid_time_w)

mean(covid_time_w$covid, na.rm = TRUE) #47.58824
median(covid_time_w$covid, na.rm = TRUE) #52 
var(covid_time_w$covid, na.rm = TRUE) #615.7271

mean(covid_time_w$mask) #10.43396
median(covid_time_w$mask) #9
var(covid_time_w$mask) #40.51959
```

```{r}
head(covid.search4$interest_by_city)
covid_city <- as_tibble(covid.search4$interest_by_city)
glimpse(covid_city)

covid_city_w <- spread(covid_city, key = keyword, value = hits)
head(covid_city_w)
maskbycity <-  covid_city_w %>%
  group_by(location) %>%
  summarize(mask = sum(mask)) %>%
  arrange(desc(mask))

na.omit(maskbycity)
#Top 5 cities for 'mask' search results are Minier, Oakdale, Bushnell, Keyesport, Forsyth
```

#Google Trends + ACS

Once you have a Census access key, save it as a text file, then read this key in the cs_key object. We
will use this object in all following API queries. 

```{r}
cs_key <- read_file("~/Downloads/census-key.txt")
```

In the following, we request basic socio-demographic information (population, median age,
median household income, income per capita) for cities and villages in the state of Illinois.
```{r}
acs_il <- getCensus(name = "acs/acs5",
                    vintage = 2020,
                    vars = c("NAME",
                             "B01001_001E",
                             "B06002_001E",
                             "B19013_001E",
                             "B19301_001E"),
                    region = "place:*",
                    regionin = "state:17",
                    key = cs_key)
head(acs_il)
```

Convert values that represent missings to NAs.
```{r}
acs_il[acs_il == -666666666] <- NA
```

Now, it might be useful to rename the socio-demographic variables (B01001_001E etc.) in our
data set and assign more meaningful names.
```{r}
acs_il <-
  acs_il %>%
  rename(pop = B01001_001E,
         age = B06002_001E,
         hh_income = B19013_001E,
         income = B19301_001E)
```

Answer the following questions with the “crime” and “loans” Google trends data and the ACS
data.

• First, check how many cities don’t appear in both data sets, i.e. cannot be matched.
Then, create a new data set by joining the Google Trends and the ACS data. Keep only
cities that appear in both data sets.
```{r}
head(res_city)
head(acs_il)

#create location variable that removes the type of location and state name from the label
acs_il$location <- gsub("\\s\\S+,.+$", "", acs_il$NAME)

anti_join(res_city_w, acs_il, by="location")
anti_join(acs_il, res_city_w, by="location")
```


``` {r}
GT_ACS1 <- res_city_w %>% inner_join(acs_il, by="location")
head(GT_ACS1)
```

• Compute the mean of the search popularity for both keywords for cities that have an
above average median household income and for those that have an below average median
household income. When building your pipe, start with creating the grouping variable
and then proceed with the remaining tasks. What conclusions might you draw from
this?

```{r}
#calculate average median income
median_income <- median(GT_ACS1$hh_income, na.rm=T)

#note that the median income is not included here
GT_ACS1$income2[GT_ACS1$hh_income < median_income] <- "Below avg median HH income"
GT_ACS1$income2[GT_ACS1$hh_income > median_income] <- "Above avg median HH income"
table(GT_ACS1$income2)

GT_ACS1 %>% summarise(crime_hits = mean(crime, na.rm=T),
                                            loans_hits = mean(loans, na.rm=T))
GT_ACS1 %>% group_by(income2) %>% summarise(crime_hits = mean(crime, na.rm=T),
                                            loans_hits = mean(loans, na.rm=T))
```

In general, there are more hits for "loans" than there are for "crime" in Illinois cities. This is true for cities with both above and below average median household incomes. However, people in cities with a below average income appear to be searching for "crime" more than those in cities with an above average income, while people in cities with above average incomes appear to be searching for "loans" more than those in cities with below average incomes. Additionally, in cities with above average median household incomes, the difference between the search volume for loans and for crime is greater than in cities with below average median household incomes.

• Is there a relationship between the median household income and the search popularity of
the Google trends terms? Describe the relationship and use a scatterplot with qplot().
```{r}
qplot(hh_income, crime, data=GT_ACS1) + geom_smooth(method='lm') + xlab("Median Household Income") +
  ylab("Search Hits (relative)") + ggtitle("Search Hits for 'Crime' by City Median Household Income")
qplot(hh_income, loans, data=GT_ACS1) + geom_smooth(method='lm') + xlab("Median Household Income") +
  ylab("Search Hits (relative)") + ggtitle("Search Hits for 'Loans' by City Median Household Income")
```

As household income increases, search volume for "crime" appears to decrease slightly. By contrast, as household income increases, search volune for "loans" increases.

###Repeat the above steps using the covid data and the ACS data.
```{r}

#merge data
GT_ACS2 <- covid_city_w %>% inner_join(acs_il, by="location")
head(GT_ACS2)

#calculate average median income
median_income <- median(GT_ACS2$hh_income, na.rm=T)

#note that the median income is not included here
GT_ACS2$income2[GT_ACS2$hh_income < median_income] <- "Below avg median HH income"
GT_ACS2$income2[GT_ACS2$hh_income > median_income] <- "Above avg median HH income"
table(GT_ACS2$income2)

GT_ACS2 %>% summarise(covid_hits = mean(covid, na.rm=T),
                                            mask_hits = mean(mask, na.rm=T))
GT_ACS2 %>% group_by(income2) %>% summarise(covid_hits = mean(covid, na.rm=T),
                                            mask_hits = mean(mask, na.rm=T))

qplot(hh_income, covid, data=GT_ACS2) + geom_smooth(method='lm') + xlab("Median Household Income") +
  ylab("Search Hits (relative)") + ggtitle("Search Hits for 'Covid' by City Median Household Income")
qplot(hh_income, mask, data=GT_ACS2) + geom_smooth(method='lm') + xlab("Median Household Income") +
  ylab("Search Hits (relative)") + ggtitle("Search Hits for 'Mask' by City Median Household Income")

```

The search volume is quite similar between "covid" and "mask" in Illinois cities. Looking by income, there is little difference in search volume for "covid" by household income, while cities with a below average median household income are slightly more likely to have searched for "mask" than cities with an above average median household income.

In looking deeper at search trends for "covid," it appears that as median household income increases, there is a slight increase in search volume for "covid." The opposite is true for "mask," where as median household income increases, there is a very slight decrease in search volume for "mask."